---
title: "Measles Outbreak"
subtitle: "Presentation"
author: "Harry Ramirez"
date: "June 11, 2019"
output: 
  ioslides_presentation: 
    keep_md: yes
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<style>
.myClass1 {
    line-height: 16pt;
    font-size: 14pt;
}
.myClass2 {
    line-height: 14pt;
    font-size: 12pt;
}
.myClass3 {
    line-height: 18pt;
    font-size: 16pt;
}
.myClass4 {
    line-height: 12pt;
    font-size: 10pt;
}

h3 {
    color: #515151;
}

</style>

## Definitions

### Outbreak

<div class='myClass3'>
According to the [CDC](https://www.cdc.gov/csels/dsepd/ss1978/lesson1/section11.html), **Epidemic** refers to an increase, often sudden, in the number of cases of a disease above what is normally expected in that population in that area. **Outbreak** carries the same definition of epidemic, but is often used for a more limited geographic area.
</div>
### Measles (Rubeola)

<div class='myClass3'>
[Measles](https://www.cdc.gov/vaccines/pubs/surv-manual/chpt07-measles.html) is an acute viral illness caused by a virus in the family paramyxovirus, genus Morbillivirus. Measles is characterized by a prodrome of fever (as high as 105°F) and malaise, cough, coryza, and conjunctivitis, followed by a maculopapular rash. The rash spreads from head to trunk to lower extremities. Measles is usually a mild or moderately severe illness. However, measles can result in complications such as pneumonia, encephalitis, and death. 
</div>

## Methodology

1. Measles data from January 1st, 2010 to May 30, 2019 was extracted from the CDC's [webpage](https://www.cdc.gov/measles/cases-outbreaks.html).

2. R code was developed to load, summarize and plot the case data.

3. Conclusions were derived from definitions, summary statistics and graphs.

4. R Markdown and Plotly were used to create the final presentation.


## R Code - Part 1

```{r echo = TRUE, message=F, warning=F, class.source='myClass2', comment=""}
library(plotly)
library(dplyr)
library(formattable)

# Load data
measles <- read.csv("MeaslesCases_5_30_2019.csv")
head(measles, 3)
```
```{r echo = TRUE, class.source='myClass2'}
# Calc summary statistics 
measles$Rate <- measles$Cases / measles$Months
casesMedian <- median(measles$Cases)
rateMedian <- median(measles$Rate)

graphTitle <- paste0("Measles Reported Cases<br>",
                     "<span style=\"font-size: 85%;",
                     "color: #144e75\">",
                     "Jan 1st, 2010 to May 30, 2019</span>")
```

## R Code - Part 2

```{r echo = TRUE, class.source='myClass2'}
# Create bar graph
p <- plot_ly(data = measles,
             x = ~Year,
             y = ~Cases,
             type = 'bar',
             name = 'Year Cases') %>%
    add_trace(y = ~Rate,
              name = 'Monthly Rate') %>%
    add_lines(y = casesMedian,
              x = ~c(min(Year)-.5, Year, max(Year)+.5),
              line = list(color = '#144E75',
                          dash = 'dash',
                          width = 2),
              name = "Case Median") %>% 
    add_lines(y = rateMedian,
              x = ~c(min(Year)-.5, Year, max(Year)+.5),
              line = list(color = '#144E75',
                          dash = 'dot',
                          width = 2),
              name = "Rate Median") %>% 
    layout(title = list(text = graphTitle,
           font = list(size=20)),
           margin = list(l = 10, r = 10, b = 50, t = 120, pad = 4),
           yaxis = list(title = 'Case Count'),
           xaxis = list(tickangle = 40),
           barmode = 'group')
```

## US Measles Incidence `r first(measles$Year)` - `r last(measles$Year)`

```{r echo = FALSE, out.width='100%'}
p
````

<br style="magin_top: 50px;">
<p class="myClass4">\* Data extracted from the [CDC Website](https://www.cdc.gov/measles/cases-outbreaks.html).</p>
<p class="myClass4">** Use zoom and pan controls to view details.</p>

## Conclusions

<div class='myClass3'>
- The total cases reported for the period between January 1st, `r first(measles$Year)` and May 30, `r last(measles$Year)` have a median of `r casesMedian` cases per year.

- The total cases reported from January 1st to May 30, `r last(measles$Year)` were `r last(measles$Cases)`, that represents a **`r digits(last(measles$Cases) / casesMedian, 2)` fold** the median cases per year.

- The median of the monthly cases rate for the last `r nrow(measles)` years was `r digits(rateMedian, 2)` cases per month. 

- The monthly rate of cases reported from January 1st to May 30, `r last(measles$Year)` was `r last(measles$Rate)` cases per month, that represents a **`r digits(last(measles$Rate) / rateMedian, 2)` fold** the median cases per month for the `r nrow(measles)` year period.

- The data shows a significant increase from the normally reported cases for the first five months of 2019, that increase can be categorize as an outbreak by CDC's definition.
</div>

<br style="margin-top: 70px">

#### **Thanks.**
