---
title: "Measles Outbreak"
subtitle: "Presentation"
author: "Harry Ramirez"
date: "June 11, 2019"
output: ioslides_presentation
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

<style>
.myClass1 {
    line-height: 16pt;
    font-size: 14pt;
}
.myClass2 {
    line-height: 14pt;
    font-size: 12pt;
}
</style>

## R Code - Part 1

```{r echo = TRUE, message=F, warning=F, class.source='myClass1', comment=""}
library(plotly)
library(dplyr)
library(formattable)

# Load data
measles <- read.csv("MeaslesCases_5_30_2019.csv")
head(measles, 3)
```
```{r echo = TRUE, class.source='myClass1'}
measles$Monthly <- measles$Cases / measles$Months
graphTitle <- paste0("US Measles Cases<br>",
                     "<span style=\"font-size: 85%;",
                     "color: #144e75\">",
                     "** Up to May 30, 2019</span>")
```

## R Code - Part 2

```{r echo = TRUE, class.source='myClass2'}
# Create bar graph
p <- plot_ly(data = measles,
             x = ~Year,
             y = ~Cases,
             type = 'bar',
             name = 'Year Cases') %>%
    add_trace(y = ~Monthly,
              name = 'Monthly Cases') %>%
    add_lines(y = ~median(Cases),
              x = ~c(min(Year)-.5, Year, max(Year)+.5),
              line = list(color = '#144E75',
                          dash = 'dash',
                          width = 2),
              name = "Year Median") %>% 
    add_lines(y = ~median(Monthly),
              x = ~c(min(Year)-.5, Year, max(Year)+.5),
              line = list(color = '#144E75',
                          width = 2),
              name = "Monthly Median") %>% 
    layout(title = list(text = graphTitle,
           font = list(size=20)),
           margin = list(l = 10, r = 10, b = 50, t = 110, pad = 4),
           yaxis = list(title = 'Cases Count'),
           xaxis = list(tickangle = 45),
           barmode = 'group')
```

## US Measles Outbreak - 2019

```{r echo = FALSE, out.width='100%'}
p
````

<br style="magin_top: 50px;">
<font size="1">Data extracted from the [CDC Website](https://www.cdc.gov/measles/cases-outbreaks.html).</font>

## Conclusions

- Up to May 30, the monthly rate of Measles cases during 2019 is `r last(measles$Monthly)`. That's *`r percent(last(measles$Monthly) / median(measles$Monthly), 1)`* the last `r nrow(measles)` years monthly median.

- With only five month of data, 2019's up to date cases are `r last(measles$Cases)`. That's *`r percent(last(measles$Cases) / median(measles$Cases))`* the last `r nrow(measles)` years median for hole years.

- We have an Outbreak.

<br style="margin-top: 100px">

### Thanks.
